---
title: "Import site data"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)

library(dplyr)
library(ggplot2)
library(readxl)
library(stringr)

# For spatial stuff
library(ggmap)
library(sf)

# Default graph theme - white background
theme_set( theme_bw() )

```


All data are in the Excel workbook `Eli_field_data14thSEP.xlsx`.

```{r}

ExcelPath <- here("data_raw", "Eli_field_data30thMAY.xlsx")

```


## Sites

Read data from worksheet `site_info`

```{r}

DAT.sites <- read_excel(ExcelPath, sheet = "site_info", guess_max = 113) %>%
  
  select(siteid = site_id,
         vegtype = f_type,
         fireclass = ffreq,
         firecount = fire_num,
         droughtclass = dsev,
         spei = SPEI,
         easting,
         northing,
         comb,
         slope,
         aspect,
         elevation = elev,
         rock,
         description = desc) %>%
  
  # Get rid of blank records (no site id)
  filter(!is.na(siteid)) %>%
  
  # Factor variables for fire, drought and veg
  mutate(fireclass = factor(fireclass, levels = 1:2, labels = c("low", "high")),
         droughtclass = factor(droughtclass, levels = 1:2, labels = c("low", "high")),
         vegtype = factor(vegtype)) %>%
  
  # First two characters of site ID can be used to put sites into spatial clusters
  # (might be useful for modelling)
  mutate(sitegroup = factor(str_sub(siteid, 1, 2)))

```


### Data checks and summaries

Fire count against fire class.

```{r}

with(DAT.sites, table(firecount, fireclass))

```

**TODO** One 'high' fire frequency site has only two fires.

```{r}

#DAT.sites %>% 
 # filter(fireclass == "high", firecount == 2) %>%
  #select(siteid)

```


SPEI values against drought class.

```{r}

dat <- filter(DAT.sites, !is.na(spei))

ggplot(data = dat) +
  geom_jitter(aes(x = droughtclass, y = spei),
              width = 0.2, height = 0.05)

```

**TODO** The 'low' drought sites have two equal-sized clusters of SPEI values, while the 'high' drought sites have a single cluster with a narrow range of values. It might be useful to define a second drought class variable to use for checking model residuals.

```{r}

DAT.sites <- DAT.sites %>%
  
  mutate(droughtclass2 = factor(
    ifelse(spei < -1.5, "high", ifelse(spei < -1.0, "med", "low")),
    levels = c("low", "med", "high")) ) %>%
  
  select(siteid, sitegroup, vegtype, 
         fireclass, firecount, 
         droughtclass, droughtclass2, 
         everything())

```


Site coordinates are zone 56 MGA. Check for any missing coordinates.

```{r}

x <- DAT.sites %>%
  filter(is.na(easting) | is.na(northing)) %>%
  select(siteid)

if (nrow(x) == 0) {
  cat("No missing coordinates")
} else {
  cat("Sites with missing coordinates:\n", x$siteid)
}

```

Graph coodinates against index to check for odd patterns or any values outside the expected range.

```{r}

plot(DAT.sites$easting)

```

```{r}

plot(DAT.sites$northing)

```

Plot site locations against a base map as a visual check.

```{r fig.width=8, fig.height=8}

dat.mga <- DAT.sites %>%
  st_as_sf(coords = c("easting", "northing"))

# Set coord ref system using the EPSG code for MGA Zone 56 (see spatialreference.org)
st_crs(dat.mga) <- 28356

# Convert to lat-lon GDA94 (= WGS84, EPSG:4326)
dat.latlon <- st_transform(dat.mga, 4326)

# Mid-point and zoom (by trial and error) for base map
x <- st_coordinates(dat.latlon)
mid <- c(lon = mean(range(x[,1])), lat = mean(range(x[,2])))
z <- 8

# Background map
suppressMessages(
  basemap <- ggmap::get_map(source = "google", 
                            location = mid, 
                            zoom = z,
                            maptype = "roadmap")
)

ggmap(basemap) +
  geom_sf(data = dat.latlon, aes(colour = sitegroup), 
          inherit.aes = FALSE, 
          size = 2, show.legend = "point") +
  
  scale_color_brewer(palette = "Set1")

```


Save data to file.

```{r}

save(DAT.sites, file = here("data", "sites.RData") )

```

